---
title: "Plotting stuff in space"
author: "David L Miller and Théo Michelot"
output:
  xaringan::moon_reader:
    nature:
      highlightStyle: github
      highlightLines: true
      slideNumberFormat: ""
      ratio: "16:9"
    seal: false
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML"
css: custom.css
---


```{r setup, include=FALSE}
library(knitr)
library(viridis)
opts_chunk$set(cache=TRUE, message=FALSE, warning=FALSE)

# some useful libraries
library(RColorBrewer)
library(ggplot2)
```

class: title-slide, inverse, center

# Plotting stuff in space

## Good practice in information conveyance
<br/>
<br/>
### David L Miller and Théo Michelot

<div style="position: absolute; bottom: 15px; vertical-align: center; right: 10px;">
<h3>CREEM Crackers 29 March 2019</h3>
</div>

---

# Setting some bounds

- Want to plot some continuous phenomena in space
  - probability of presence, utilization, abundance etc
- Values are important (not just "background")


---

# WE ARE NOT GOING TO TALK ABOUT:

- colour schemes
- projections
- whether `ggplot2` is better than `base` graphics
- how ugly Dave's plots are in general

---

```{r, echo=FALSE}
load("data/afttdat.RData")
tp <- subset(plotdat, model=="tp")
pf <- subset(plotdat, model=="pf")
```

# Some data for examples

- Predictions of sperm whale abundance over the US AFTT
- A few issues:
  - 2 sets of model results, need to facet
  - 1 model "tp" has huge range

```{r}
range(tp$p)
range(pf$p)
```

---

# Basic plot

```{r, fig.width=10, fig.height=4}
ggplot(plotdat, aes(x=x,y=y, fill=p)) +
  geom_tile() +
  coord_equal(xlim=range(plotdat$x), ylim=range(plotdat$y)) +
  theme_minimal() +
  us +
  facet_wrap(~model)
```

---

# Just "pf"

.pull-left[
```{r, echo=FALSE}
basic_pf <- ggplot(pf, aes(x=x,y=y, fill=p)) +
  geom_tile() +
  coord_equal(xlim=range(plotdat$x), ylim=range(plotdat$y)) +
  theme_minimal() +
  us
print(basic_pf)
```
]

.pull-right[
```{r, eval=FALSE}
ggplot(pf, aes(x=x,y=y, fill=p)) +
  geom_tile() +
  coord_equal(xlim=range(plotdat$x), ylim=range(plotdat$y)) +
  theme_minimal() +
  us
```
]



---

# Just "tp"

.pull-left[
```{r, echo=FALSE}
basic_tp <- ggplot(tp, aes(x=x,y=y, fill=p)) +
  geom_tile() +
  coord_equal(xlim=range(plotdat$x), ylim=range(plotdat$y)) +
  theme_minimal() +
  us
print(basic_tp)
```
]

.pull-right[
```{r, eval=FALSE}
ggplot(tp, aes(x=x,y=y, fill=p)) +
  geom_tile() +
  coord_equal(xlim=range(plotdat$x), ylim=range(plotdat$y)) +
  theme_minimal() +
  us
```
]


---

# First let's focus on "tp"

Main issue: what to do with wide range of values?

```{r}
basic_tp + scale_fill_viridis_c(option="E")
```

---

# log scale?

```{r}
basic_tp + scale_fill_viridis_c(option="E", trans="log")
```

---

# To contour or not to contour?

( This is for pf )

```{r}
basic_pf + scale_fill_viridis_c(option="E") +
  geom_contour(aes(z=p))
```

---


# That default looks awful

( in `ggplot2` you don't get contour labels? )

```{r}
basic_pf + scale_fill_viridis_c(option="E") +
  geom_contour(aes(z=p), binwidth=1)
```



---

# Are continuous legends actually useless?

```{r}
basic_tp + scale_fill_viridis_c(option="E")
```

---



